---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# burgle

<!-- badges: start -->
<!-- badges: end -->

The goal of burgle is to "steal" only the necessary parts of model objects.

## Installation

You can install the development version of burgle like so:

``` {r, eval = FALSE}
devtools::install_github("ClevelandClinicQHS/burgle")
```

## Pros of `burgle`

 1. The reduction in size can save on memory and storage space. Something that is sometimes in great demand since R works with local active memory.
 2. The removal of data from the model objects can allow them to be shared freely, if there are data sharing concerns or requirements.
 3. A streamline to simulate response for parameter uncertainty and probabilistic sampling


## Linear Model Example

```{r example}

set.seed(287453)
library(burgle)
fit <- lm(Sepal.Length ~., data = iris)
bfit <- burgle(fit)
object.size(fit)
object.size(bfit)

as.numeric(object.size(bfit)/object.size(fit))*100
```

Our `burgle_lm` is roughly `r paste0(round(as.numeric(object.size(bfit)/object.size(fit))*100,2), "%")` the size of the original `lm` object, the iris dataset has `r nrow(iris)` observations and `r ncol(iris)` columns.

Another example is the using the `nycflights13::flights` dataset. 

```{r larger_example}
fit2 <- lm(arr_delay ~ as.factor(month) + dep_delay + origin + distance + hour, data = nycflights13::flights)
b_fit2 <- burgle(fit2)

as.numeric(object.size(b_fit2)/object.size(fit2))*100
```

Our `burgle_lm` is roughly `r paste0(round(as.numeric(object.size(b_fit2)/object.size(fit2))*100,2), "%")` the size of the original `lm` object. This dataset has `r nrow(nycflights13::flights)` observations and our model has used 5 of the `r ncol(nycflights13::flights)` columns as predictors.

## Prediction

The new predict methods for our `burgle` objects allow for one to easily predict using `newdata`. 
The structure is as follows:

 * The rows are indexed by the original row order given in the `newdata`
 * The columns are the different models (number of `draws` set to 1 by default)
 * If more than one simulation is done, then the simulations are items in a list per model and row observation (number of `sims` set to 1 by default)

## Predictions

If one wants to predict using the original model simply set `original = TRUE`.

Depending on the model object there are different types of predictions. By default it will return the linear predictor `(lp)`. If one wants to see the response, which makes more sense when looking at `glm` and survival models, one can set the `type = "response"`. The `se = FALSE` is an argument on whether to include the standard error of the model when simluating responses. `TRUE` means to use the model standard error. We recommend setting it to FALSE when doing more than one simulation or when setting `type = "response"`.
```{r predictions}
predict(bfit, newdata = head(iris), original = TRUE, draws = 1, se = FALSE, type = "lp")
predict(bfit, newdata = head(iris), original = FALSE, draws = 5, type = "lp")
## These two should be similar
predict(bfit, newdata = head(iris), original = FALSE, draws = 5, se = TRUE, type = "response")
predict(bfit, newdata = head(iris), original = FALSE, draws = 5, sims = 5, se = TRUE, type = "response")
```


## Generalized Linear Model

The framework should work with all `glm` family options, just the binomial one is demonstrated below.
```{r competiting risk}
b_glm <- burgle(glm(I(Species == "versicolor") ~ ., family = "binomial", data = iris))

predict(b_glm, head(iris), original = FALSE, se = TRUE, draws = 5, type = "lp")
predict(b_glm, head(iris), original = FALSE, se = TRUE, draws = 5, type = "response")
```

## Cox Proporiontal Hazards Model
```{r}
library(survival)
lung <- survival::lung
lung$status <- lung$status - 1

cox <- coxph(Surv(time, status) ~ age + sex + ph.ecog + ph.karno + pat.karno, data = lung)
cox_sm <- coxph(Surv(time, status) ~ age + sex + ph.ecog + ph.karno + pat.karno, data = lung, x = FALSE, y = FALSE)

b_cox <- burgle(cox)
object.size(cox)
object.size(cox_sm)
object.size(b_cox)

as.numeric(object.size(b_cox)/object.size(cox))*100
as.numeric(object.size(b_cox)/object.size(cox_sm))*100
```

Our `burgle_coxph` model is `r paste0(round(as.numeric(object.size(b_cox)/object.size(cox_sm))*100,2), "%")` the size of the original Cox proportional hazards model even after setting `x=FALSE` and `y= FALSE`. The lung dataset has `r nrow(lung)` observations.

One way to further reduce the size of the `burgle_coxph` is to reduce the number of unique time points in the data, since the it contains the baseline hazard of the model. The lung dataset as `r length(unique(lung$time))` unique values. If we were to round these to the nearest 14 days that would reduce the number of timepoints to `r length(unique(plyr::round_any(lung$time, 14)))`.

```{r}
lung$time2 <- plyr::round_any(lung$time, 14)
cox2 <- coxph(Surv(time2, status) ~ age + sex + ph.ecog + ph.karno + pat.karno, data = lung, x = TRUE, y = FALSE)
b_cox2 <- burgle(cox2)

object.size(cox2)
object.size(b_cox2)
as.numeric(object.size(b_cox2)/object.size(cox2))*100
```

This reduce the size to `r paste0(round(as.numeric(object.size(b_cox2)/object.size(cox2))*100,2), "%")` of the original `coxph` object.

## Survival predictions

Predictions are a slightly different structure if `type = "response"` or `"risk"`, since a time point is required. If `type = "response"` the returned results is a simulated 1 or 0 if the event has been experienced or not at a given time point. 
The structure is as follows:
 * The rows are indexed by the original row order given in `newdata`
 * The columns are the different time points at which the risk is calculated
 * The different elements of the lists are the different models (`draws`)
 * If more than one simulation is done, then the different simulations are returned as list within the list element for each model (see below for an example example)
```{r}
predict(b_cox, newdata = head(lung), original = TRUE, draws = 1, type = "lp")
predict(b_cox, newdata = head(lung), original = TRUE, draws = 1, type = "risk", times = 500)
predict(b_cox, newdata = head(lung), original = TRUE, draws = 1, type = "risk", times = c(500, 1000))
predict(b_cox, newdata = head(lung), original = TRUE, draws = 1, type = "response", times = c(500, 1000))
predict(b_cox, newdata = head(lung), original = FALSE, draws = 5, sims = 2, type = "risk", times = c(500, 1000))

```

## Larger Example
```{r}
## The original model at time 500
predict(b_cox, newdata = head(lung), original = TRUE, draws = 1, type = "risk", times = c(500))

## Doing 1000 simulations from the original model and calculating the death rate
a0 <- predict(b_cox, newdata = head(lung), original = TRUE, draws = 1, sims = 1000, type = "response", times = c(500)) |> purrr::list_flatten() |> Reduce(f = cbind, x = _) |> apply(1, mean)
a0

## Average survival death rate based on 1000 different models
a1 <- predict(b_cox, newdata = head(lung), original = FALSE, draws = 1000, type = "response", times = c(500)) |> purrr::list_flatten() |> Reduce(f = cbind, x = _) |> apply(1, mean)

a1

## Average survival rate based on 100 simlutions for each of the 1000 models
a2 <- predict(b_cox, newdata = head(lung), original = FALSE, draws = 1000, sims = 100, type = "response", times = c(500))

### Average death per model
a3 <- lapply(a2, function(x) apply(Reduce(cbind, x), 1, mean))

a4 <- Reduce(rbind, a3)

apply(a4, 2, median)
```


This structed has also been implemented for `riskRegression::CSC objects` and the plan is to incorportate `rstan` objects, and an overall default method and which will require a mean and covariance matrix as inputs.
